Modelling BCG vaccination in the UK

### 2019-07-04 rmds/08-model-development.Rmd
shortened table and figure captions in lists

```{r latency-flow-diag, out.width = "200px", fig.cap = "Flow diagrams of a.) the serial latency assumption and b.) the parallel latency assumption. The flow diagrams contain the following compartments; Susceptible ($S$), high risk latent ($H$), low risk latent ($L$), and infected ($I$). Solid arrows represent transition rates. Note that in both models repeated transmission to low risk latents is possible. This allows low risk latent cases to become high risk latent cases. For some varients of the parallel latency assumption, where it is assumed being high risk is inherent to individuals, this may not be appropriate."}
knitr::include_graphics(file.path(resources_path, "resources", "model_diagrams", "latency_flow.png"))
```

@@

```{r model-flow-diag, out.width = "400px", fig.cap = "Flow diagram for the dynamic TB disease model with demographics and vaccination described in the text. The TB model contains the following compartments; Susceptible ($S$), high risk latent ($H$), low risk latent ($L$), active cases with pulmonary TB ($P$), active TB cases with extra-pulmonary TB only ($E$), pulmonary cases on treatment ($T_P$), and extra-pulmonary cases on treatment ($T_E$). The vaccinated ($v$) and unvaccinated ($u$) populations are represented by $k$, such that $k = u,v$. Age stratification is represented by $a$ (where $a = 1, 2, ...,11$) in the disease model and the $0,\ 1,\ 2,\ 3$ subscripts in the demographic model. Five year age groups are modelled (i.e $0-4$, $5-9$, $10-14$, ...) up to 49 years old, with a single age group for those aged 50-69 years old and those aged 70-89 years old. Individuals aged $90+$ are not explicitly modelled. In the demographic and vaccination model the A compartment represents the demographic processes modelled in all population compartments except for the vacinated and unvaccinated susceptible populations. Solid arrows represent transition rates within the modelled populations and dashed arrows represent transition rates into, or out of the modelled populations (i.e birth and death processes)."}
knitr::include_graphics(file.path(resources_path, "resources", "model_diagrams", "tb_model_flow.png"))
```

@@

```{r disease-model}
disease_params_table %>% 
  select(Parameter, Description, Distribution, Units, Method, Type) %>% 
kable(caption = "Dynamic disease model parameters, descriptions, prior distributions, units, method used to derive the prior distribution and the type (i.e data derived, literature, assumption). All data based parameters are included. P = pulmonary TB, E = extra-pulmonary TB, v = vaccinated, i = age at vaccination, $\\mathcal{U}$ = Uniform, $\\mathcal{N}$ = Normal", booktabs = TRUE, longtable = TRUE, escape = FALSE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header", "hold_position")) %>% 
  column_spec(c(5), width = "6cm") %>% 
  column_spec(c(2), width = "4cm") %>%
  column_spec(c(3), width = "6cm") %>%
  column_spec(c(1,4,6), width = "1.5cm") %>% 
  landscape()
```

@@

```{r tb-time-treat, fig.cap = "Distribution of time to treatment (days) from the date of reported symptom onset until the date started treatment for the UK born, stratified by age group and pulmonary/extra-pulmonary TB status in the ETS system for notifications between 2000 and 2012. Age is stratified into three groups; children (0-14), adults (15-59) older adults (15-59). The time from symptom onset to starting treatment is shorter for cases with pulmonary TB cases across age groups, with younger cases starting treatment more rapidly than older cases."}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "time_to_treatment.png"))
```

@@

```{r tb-treat-time-succ, fig.cap = "Distribution of time to treatment completion in the UK born successfully treated (days), stratified by age group and pulmonary/extra-pulmonary TB status in the ETS for notifications between 2000 and 2012. Age is stratified into three groups; children (0-14), adults (15-59) older adults (15-59). There is little evidence that the time to successful treatment differs between pulmonary and extra-pulmonary cases only but older cases appear to have a high likelihood of longer treatment times."}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "time_cure.png"))
```

@@

```{r contact-tile-plot, fig.cap = "a.) Mean contact rates and the b.) normalised standard deviation (\\%) of 1000 boostrapped samples of  social contacts from the POLYMOD social contact survey using 5 year age groups up 49 years old and then a single group for 50-69 year olds. Mixing is highly assortative by age with children and young adults representing the majority of contacts. There is also evidence of mixing between children and middle age adults with older children mixing with progressivly older adults. Contact rates in older adults are highly uncertain, with the most uncertainty in mixing between older adults and young children."}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "contact_matrix.png"))
```

@@

```{r bcg-eff-est}
kable(eff_bcg, booktabs = TRUE, escape = TRUE,
      caption = "Estimates of the effectiveness of the BCG vaccine at preventing active TB disease stratified by years since vaccination. For 0-9 years since vaccination estimates were derived using Poisson regression from the MRC BCG trial and for 10-29 years since vaccination estimates were extracted from a more recent case control cohort study in the UK born vaccinated at school-age.[@Hart1972; @Mangtani2017] BCG effectiveness at preventing active TB disase used as prior distributions in the model.") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r demo-model-params, results = "asis"}
demographic_params_table %>% 
  select(Parameter, Description, Distribution, Units, Method, Type) %>% 
kable(caption = "Demographic model parameters, descriptions, prior distributions, units, method used to derive the prior distribution and the type (i.e data derived, literature, assumption). $\\mathcal{U}$ = Uniform and i = age at vaccination.", booktabs = TRUE, longtable = TRUE, escape = FALSE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header", "hold_position")) %>% 
  column_spec(c(5), width = "6cm") %>% 
  column_spec(c(2), width = "4cm") %>%
  column_spec(c(3), width = "6cm") %>%
  column_spec(c(1,4,6), width = "1.5cm") %>% 
  landscape()
```

@@

```{r age-strat-demo, fig.cap = "Distribution of the UK born population of England in 2000, 2004, 2008, and 2012. Age is grouped into 5 year age groups from 0 to 49, from 50-69, and from 70 to 89. Those aged 90+ are excluded due to low quality data. The age groups used here represent those used in the model. The figure indicates that the population has skewed older overall over the last two decades, although the proportion of young children has increased in the last 10 years."}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "england_demographics.png"))
```

@@

```{r births-england, fig.cap = "Estimated and projected live births in England from 1929 until 2101. The red line indicates estimated data and the blue line indicates projected data. Data is sourced from the Office for National Statistics (ONS)."}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "births.png"))
```

@@

```{r mortality-england, fig.cap = "3 year rolling average expected remaining lifespan stratified by age group in England from 2000 to 2014. Age is grouped into 5 year age groups from 0 to 49, from 50-69, and from 70 to 89. Those aged 90+ are excluded due to low quality data. The age groups used here represent those used in the model. Data from this figure was sourced from the Office for National Statistics age-specific mortality rate estimates."}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "mortality.png"))
```

>>>>

```{r latency-flow-diag, out.width = "200px", fig.scap = "Flow diagrams of a.) the serial latency assumption and b.) the parallel latency assumption.", fig.cap = "Flow diagrams of a.) the serial latency assumption and b.) the parallel latency assumption. The flow diagrams contain the following compartments; Susceptible ($S$), high risk latent ($H$), low risk latent ($L$), and infected ($I$). Solid arrows represent transition rates. Note that in both models repeated transmission to low risk latents is possible. This allows low risk latent cases to become high risk latent cases. For some varients of the parallel latency assumption, where it is assumed being high risk is inherent to individuals, this may not be appropriate.", out.extra = ""}
knitr::include_graphics(file.path(resources_path, "resources", "model_diagrams", "latency_flow.png"))
```

@@

```{r model-flow-diag, out.width = "400px", fig.scap = "Flow diagram for the dynamic TB disease model with demographics and vaccination described.", fig.cap = "Flow diagram for the dynamic TB disease model with demographics and vaccination described. The TB model contains the following compartments; Susceptible ($S$), high risk latent ($H$), low risk latent ($L$), active cases with pulmonary TB ($P$), active TB cases with extra-pulmonary TB only ($E$), pulmonary cases on treatment ($T_P$), and extra-pulmonary cases on treatment ($T_E$). The vaccinated ($v$) and unvaccinated ($u$) populations are represented by $k$, such that $k = u,v$. Age stratification is represented by $a$ (where $a = 1, 2, ...,11$) in the disease model and the $0,\ 1,\ 2,\ 3$ subscripts in the demographic model. Five year age groups are modelled (i.e $0-4$, $5-9$, $10-14$, ...) up to 49 years old, with a single age group for those aged 50-69 years old and those aged 70-89 years old. Individuals aged $90+$ are not explicitly modelled. In the demographic and vaccination model the A compartment represents the demographic processes modelled in all population compartments except for the vacinated and unvaccinated susceptible populations. Solid arrows represent transition rates within the modelled populations and dashed arrows represent transition rates into, or out of the modelled populations (i.e birth and death processes).", out.extra = ""}
knitr::include_graphics(file.path(resources_path, "resources", "model_diagrams", "tb_model_flow.png"))
```

@@

```{r disease-model}
disease_params_table %>% 
  select(Parameter, Description, Distribution, Units, Method, Type) %>% 
kable(
  caption.short = "Dynamic disease model parameters, descriptions, prior distributions, units, method used to derive the prior distribution and the type.", 
  caption = "Dynamic disease model parameters, descriptions, prior distributions, units, method used to derive the prior distribution and the type (i.e data derived, literature, assumption). All data based parameters are included. P = pulmonary TB, E = extra-pulmonary TB, v = vaccinated, i = age at vaccination, $\\mathcal{U}$ = Uniform, $\\mathcal{N}$ = Normal", booktabs = TRUE, longtable = TRUE, escape = FALSE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header", "hold_position")) %>% 
  column_spec(c(5), width = "6cm") %>% 
  column_spec(c(2), width = "4cm") %>%
  column_spec(c(3), width = "6cm") %>%
  column_spec(c(1,4,6), width = "1.5cm") %>% 
  landscape()
```

@@

```{r tb-time-treat, fig.scap = "Distribution of time to treatment (days) from the date of reported symptom onset until the date started treatment for the UK born, stratified by age group and pulmonary/extra-pulmonary TB status in the ETS system for notifications between 2000 and 2012.", fig.cap = "Distribution of time to treatment (days) from the date of reported symptom onset until the date started treatment for the UK born, stratified by age group and pulmonary/extra-pulmonary TB status in the ETS system for notifications between 2000 and 2012. Age is stratified into three groups; children (0-14), adults (15-59) older adults (15-59). The time from symptom onset to starting treatment is shorter for cases with pulmonary TB cases across age groups, with younger cases starting treatment more rapidly than older cases.", out.extra = ""}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "time_to_treatment.png"))
```

@@

```{r tb-treat-time-succ, fig.scap = "Distribution of time to treatment completion in the UK born successfully treated (days), stratified by age group and pulmonary/extra-pulmonary TB status in the ETS for notifications between 2000 and 2012.", fig.cap = "Distribution of time to treatment completion in the UK born successfully treated (days), stratified by age group and pulmonary/extra-pulmonary TB status in the ETS for notifications between 2000 and 2012. Age is stratified into three groups; children (0-14), adults (15-59) older adults (15-59). There is little evidence that the time to successful treatment differs between pulmonary and extra-pulmonary cases only but older cases appear to have a high likelihood of longer treatment times.", out.extra = ""}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "time_cure.png"))
```

@@

```{r contact-tile-plot, fig.scap = "a.) Mean contact rates and the b.) normalised standard deviation (\\%) of 1000 boostrapped samples of  social contacts from the POLYMOD social contact survey using 5 year age groups up 49 years old and then a single group for 50-69 year olds.", fig.cap = "a.) Mean contact rates and the b.) normalised standard deviation (\\%) of 1000 boostrapped samples of  social contacts from the POLYMOD social contact survey using 5 year age groups up 49 years old and then a single group for 50-69 year olds. Mixing is highly assortative by age with children and young adults representing the majority of contacts. There is also evidence of mixing between children and middle age adults with older children mixing with progressivly older adults. Contact rates in older adults are highly uncertain, with the most uncertainty in mixing between older adults and young children.", out.extra = ""}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "contact_matrix.png"))
```

@@

```{r bcg-eff-est}
kable(eff_bcg, booktabs = TRUE, escape = TRUE,
      caption.short = "Estimates of the effectiveness of the BCG vaccine at preventing active TB disease stratified by years since vaccination.", 
      caption = "Estimates of the effectiveness of the BCG vaccine at preventing active TB disease stratified by years since vaccination. For 0-9 years since vaccination estimates were derived using Poisson regression from the MRC BCG trial and for 10-29 years since vaccination estimates were extracted from a more recent case control cohort study in the UK born vaccinated at school-age.[@Hart1972; @Mangtani2017] BCG effectiveness at preventing active TB disase used as prior distributions in the model.") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r demo-model-params, results = "asis"}
demographic_params_table %>% 
  select(Parameter, Description, Distribution, Units, Method, Type) %>% 
kable(
  caption.short = "Demographic model parameters, descriptions, prior distributions, units, method used to derive the prior distribution and the type.", 
  caption = "Demographic model parameters, descriptions, prior distributions, units, method used to derive the prior distribution and the type (i.e data derived, literature, assumption). $\\mathcal{U}$ = Uniform and i = age at vaccination.", booktabs = TRUE, longtable = TRUE, escape = FALSE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header", "hold_position")) %>% 
  column_spec(c(5), width = "6cm") %>% 
  column_spec(c(2), width = "4cm") %>%
  column_spec(c(3), width = "6cm") %>%
  column_spec(c(1,4,6), width = "1.5cm") %>% 
  landscape()
```

@@

```{r age-strat-demo, fig.scap = "Distribution of the UK born population of England in 2000, 2004, 2008, and 2012.", fig.cap = "Distribution of the UK born population of England in 2000, 2004, 2008, and 2012. Age is grouped into 5 year age groups from 0 to 49, from 50-69, and from 70 to 89. Those aged 90+ are excluded due to low quality data. The age groups used here represent those used in the model. The figure indicates that the population has skewed older overall over the last two decades, although the proportion of young children has increased in the last 10 years.", out.extra = ""}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "england_demographics.png"))
```

@@

```{r births-england, fig.scap = "Estimated and projected live births in England from 1929 until 2101.", fig.cap = "Estimated and projected live births in England from 1929 until 2101. The red line indicates estimated data and the blue line indicates projected data. Data is sourced from the ONS.", out.extra = ""}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "births.png"))
```

@@

```{r mortality-england, fig.scap = "3 year rolling average expected remaining lifespan stratified by age group in England from 2000 to 2014.", fig.cap = "3 year rolling average expected remaining lifespan stratified by age group in England from 2000 to 2014. Age is grouped into 5 year age groups from 0 to 49, from 50-69, and from 70 to 89. Those aged 90+ are excluded due to low quality data. The age groups used here represent those used in the model. Data from this figure was sourced from the ONS age-specific mortality rate estimates.", out.extra = ""}
knitr::include_graphics(file.path(resources_path, "resources", "figure", "mortality.png"))
```

### 2019-07-04 rmds/09-model-fitting.Rmd
shortened table and figure captions in lists

```{r measurement-model}
measurement_params_table %>% 
  select(Parameter, Description, Distribution, Units, Method, Type) %>% 
kable(caption = "Measurement model parameters, descriptions, prior distributions, units, method used to derive the prior distribution and the type (i.e data derived, literature, assumption). $\\mathcal{U}$ = Uniform", booktabs = TRUE, longtable = TRUE, escape = FALSE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(5), width = "6cm") %>% 
  column_spec(c(2), width = "4cm") %>%
  column_spec(c(3), width = "6cm") %>%
  column_spec(c(1,4,6), width = "1.5cm") %>% 
  landscape()
```

>>>>

```{r measurement-model}
measurement_params_table %>% 
  select(Parameter, Description, Distribution, Units, Method, Type) %>% 
kable(
  caption.short = "Measurement model parameters, descriptions, prior distributions, units, method used to derive the prior distribution and the type.",
  caption = "Measurement model parameters, descriptions, prior distributions, units, method used to derive the prior distribution and the type (i.e data derived, literature, assumption). $\\mathcal{U}$ = Uniform", booktabs = TRUE, longtable = TRUE, escape = FALSE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(5), width = "6cm") %>% 
  column_spec(c(2), width = "4cm") %>%
  column_spec(c(3), width = "6cm") %>%
  column_spec(c(1,4,6), width = "1.5cm") %>% 
  landscape()
```

### 2019-07-04 rmds/index.Rmd
shortened table and figure captions in lists

---
author: 'Samuel Abbott'
date: 'May 2019'
division: 'Bristol Medical School: Population Health Sciences'
advisor: 'Ellen Brooks Pollock'
altadvisor: 'Hannah Christensen'
department: 'Faculty of Health Sciences'
title: 'Modelling BCG vaccination in the UK'
subtitle: 'What is the impact of changing policy?'
wordcount: "41,218"
knit: bookdown::render_book
site: bookdown::bookdown_site
always_allow_html: yes
output: 
  #thesisdown::thesis_pdf: default
  thesisdown::thesis_gitbook: default
  #thesisdown::thesis_word: default
  #thesisdown::thesis_epub: default
abstract: >
  `r if(knitr:::is_latex_output()) paste(readLines("00-abstract.Rmd"), collapse = ' ')`
acknowledgements: >
 `r if(knitr:::is_latex_output()) paste(readLines("00-acknowledgements.Rmd"), collapse = ' ')`
declaration: >
  `r if(knitr:::is_latex_output()) paste(readLines("00-declaration.Rmd"), collapse = ' ')`
contribution: >
  `r if(knitr:::is_latex_output()) paste(readLines("00-contribution.Rmd"), collapse = ' ')`
bibliography: bib/thesis.bib
#Download your specific bibliography database file and refer to it in the line above.
csl: csl/bmj.csl
#Download your specific csl file and refer to it in the line above.
lot: true
lof: true
space_between_paragraphs: true
#Delete the # at the beginning of the previous line if you'd like
#to have a blank new line between each paragraph
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage[table]{xcolor}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
---

>>>>

---
author: 'Samuel Abbott'
date: 'August 2019'
division: 'Bristol Medical School: Population Health Sciences'
advisor: 'Ellen Brooks Pollock'
altadvisor: 'Hannah Christensen'
department: 'Faculty of Health Sciences'
title: 'Modelling BCG vaccination in the UK'
subtitle: 'What is the impact of changing policy?'
wordcount: "41,218"
knit: bookdown::render_book
site: bookdown::bookdown_site
always_allow_html: yes
output: 
  thesisdown::thesis_pdf: default
  #thesisdown::thesis_gitbook: default
  #thesisdown::thesis_word: default
  #thesisdown::thesis_epub: default
abstract: >
  `r if(knitr:::is_latex_output()) paste(readLines("00-abstract.Rmd"), collapse = ' ')`
acknowledgements: >
 `r if(knitr:::is_latex_output()) paste(readLines("00-acknowledgements.Rmd"), collapse = ' ')`
declaration: >
  `r if(knitr:::is_latex_output()) paste(readLines("00-declaration.Rmd"), collapse = ' ')`
contribution: >
  `r if(knitr:::is_latex_output()) paste(readLines("00-contribution.Rmd"), collapse = ' ')`
acronyms: >
  `r if(knitr:::is_latex_output()) paste(readLines("00-acronyms.Rmd"), collapse = ' ')`
bibliography: bib/thesis.bib
#Download your specific bibliography database file and refer to it in the line above.
csl: csl/bmj.csl
#Download your specific csl file and refer to it in the line above.
lot: true
lof: true
space_between_paragraphs: true
#Delete the # at the beginning of the previous line if you'd like
#to have a blank new line between each paragraph
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage[table]{xcolor}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
---
